---
title: "Biol 307 Project"
author: "Amyre"
date: "2025-03-31"
output: pdf_document
---

## Set Up 

***load libraries***

```{r}
library(ggplot2)
library(dplyr)
```

***read in data***

```{r}
quadrat_data<- read.csv("./data/QuadratData.csv")
tree_data<- read.csv("./data/TreeMeasurements.csv")
```

***calculate important metrics***

```{r}
#add diamter over 7 means to main dataframe
mean_diameter<- tree_data %>%
  group_by(Quadrat_ID) %>%
  summarise(mean_measurement = mean(Tree_Measurement))

quadrat_data$MeanDiameter<- mean_diameter$mean_measurement

#relative abundance of large trees (trees over 7cm diameter)
quadrat_data$rel_large<- quadrat_data$Over_7/quadrat_data$Total
```

***format categorical data***

```{r}
quadrat_data$Quadrat_ID<- as.factor(quadrat_data$Quadrat_ID)

quadrat_data$Trail<- trimws(quadrat_data$Trail)
quadrat_data$Trail<- as.factor(quadrat_data$Trail)

quadrat_data$Dist_Arroyo<- as.factor(quadrat_data$Dist_Arroyo)
```

***subset near and far data***

```{r}
near_data<- subset(quadrat_data, Dist_Arroyo=="Near")

far_data<- subset(quadrat_data, Dist_Arroyo=="Far")
```

***check for normality***

```{r}
shapiro.test(quadrat_data$Total)
shapiro.test(quadrat_data$rel_large)
shapiro.test(quadrat_data$MeanDiameter)

```

## Compare Total Number of Trees 

***visualize***

```{r}
Figure3<- ggplot(quadrat_data, aes(Dist_Arroyo, Total, fill = Dist_Arroyo))+
  geom_boxplot()+
  labs(x="Distance from Arroyo", y="Number of Trees")+
  theme(legend.position = "none")
```

***independent t-test***

```{r}
t_test_result1 <- t.test(near_data$Total, far_data$Total)
print(t_test_result1)
```

## Comparing Relative Abundance of Large Trees

***visualize***

```{r}
Figure4<- ggplot(quadrat_data, aes(Dist_Arroyo, rel_large, fill = Dist_Arroyo))+
  geom_boxplot()+
  labs(x="Distance from Arroyo", y="Proportion of Trees over 7cm DBH")+
  theme(legend.position = "none")
```

***independent t-test***

```{r}
t_test_result2 <- t.test(near_data$rel_large, far_data$rel_large)
print(t_test_result2)
```

## Comparing Mean DBH in Trees Above 7cm DBH

***visualize***

```{r}
Figure5<- ggplot(quadrat_data, aes(Dist_Arroyo, MeanDiameter, fill = Dist_Arroyo))+
  geom_boxplot()+
  labs(x="Distance from Arroyo", y="Mean DBH in Trees over 7cm DBH")+
  theme(legend.position = "none")
```

***independent t-test***

```{r}
t_test_result3 <- t.test(near_data$MeanDiameter, far_data$MeanDiameter)
print(t_test_result3)
```

## Controlling for Potenial Confounding Variables

***ANCOVA***

```{r}
ancova_model1 <- aov(Total ~ Dist_Arroyo + Trail, data = quadrat_data)
summary(ancova_model1)

ancova_model2 <- aov(rel_large ~ Dist_Arroyo + Trail, data = quadrat_data)
summary(ancova_model2)

ancova_model3 <- aov(MeanDiameter ~ Dist_Arroyo + Trail, data = quadrat_data)
summary(ancova_model3)
```

## Save plots 

```{r}
ggsave("./Figures/Fig3.png", plot = Figure3, width = 8, height = 6)
ggsave("./Figures/Fig4.png", plot = Figure4, width = 8, height = 6)
ggsave("./Figures/Fig5.png", plot = Figure5, width = 8, height = 6)

```
